```{r setup, include=FALSE}
library(later)
```


# Later

Scoped side effects and events for R.

[![Travis-CI Build Status](https://travis-ci.org/kevinushey/later.svg?branch=master)](https://travis-ci.org/kevinushey/later)

This package:

- generalizes `on.exit()` to a function `defer()`, which can attach exit
  handlers to any currently executing function, and uses that to scope side
  effects, and

- provides a simple event system, using event handlers registered
  through `on()` / `off()`, and events fired through `emit()`.

## Scoped Side Effects

The `defer()` function generalizes `on.exit()`, allowing a function to perform 
cleanup work after a parent invoking function has finished execution. For
example, the following `scope_dir(dir)` function can be used to set the
working directory for duration of a function's execution:

```{r}
scope_dir <- function(dir) {
  owd <- setwd(dir)
  defer(setwd(owd), envir = parent.frame())
}
```

The `scope_dir()` function can be called to set the working directory, and
restore the working directory when the caller is done. For example:


```{r}
in_dir <- function(dir) {
  scope_dir(dir)
  print(basename(getwd()))
}
print(basename(getwd()))
in_dir("tests")
print(basename(getwd()))
```

## Events

Register an event handler using `on("event", <handler>)`, and fire events with
`emit("event", <data>)`. Useful for long-range communication between
functions / objects.

These functions are similar to builtin `R` capabilities: `signalCondition()` 
allows you to signal a condition, and `withCallingHandlers()` can be used to
register handlers for signaled conditions (alongside errors, warnings, and
otherwise). The main difference between the system in `later` and in base `R`:

1. Rather than wrapping the executing expression in `withCallingHandlers()`, you
   can just call `on()` anywhere in a function's body, and any events emitted
   later on in any `R` code will be handled by that handler,

2. A call to `emit()` can emit _any_ kind of data object; you are not limited
   to `R` conditions,

3. A registered handler for an event won't alter control flow (ie, this event
   system doesn't have the `restarts` mechanism that `R` uses for 'long jumps'),

4. Handlers are registered in a stack (so the most recently registered handlers
   will handle an event first); and handlers can call `stop_propagation()` to
   ensure a handler deeper in the stack does not receive the event.

An example to illustrate, with a set of 'workers' that can call `emit()`
when they've completed some work, and a 'manager' that listens for the
emitted signals from these workers.

```{r}
set.seed(1)
make_worker <- function(i) {
  force(i)
  function() {
    emit("work", i)
  }
}

workers <- lapply(seq_len(5), make_worker)
manager <- function() {
  
  counter <- 0
  on("work", function(data) {
    cat(sprintf("Received data: '%s'\n", data), sep = "")
    counter <<- counter + 1
  })
  
  # Run for 2 milliseconds
  time <- Sys.time()
  while (Sys.time() - time < 0.002) {
    worker <- sample(workers, 1)[[1]]
    worker()
  }
  
  cat(sprintf("Executed %s tasks.\n", counter), sep = "")
}

manager()
```

## Inspiration & Credit

This package is heavily inspired by Jim Hester's
[withr](https://github.com/jimhester/withr) package.

The main 'trick' that enabled the generalization of `on.exit()` was provided in
a mailing list post
here, by Peter Meilstrup: https://stat.ethz.ch/pipermail/r-devel/2013-November/067874.html
